---
title: "Task 3: Santa Barbara Urchins"
author: "Julia Wilson"
date: "10/23/2021"
output: 
  html_document: 
    toc: yes
    theme: cerulean
---
### Overview
In this task we wrangle, analyze, and visualize data about red and purple sea urchins recorded by Santa Barbara Coastal Long Term Ecological Research program divers. 

**Data Citation:** Reed, LTER: Reef:Long-term experiment:Kelop removal: Urchin size frequency distribution ver 20. Environmental Data Initiative. https://doi.org/10.6073/pasta/5a1e9ef03aa47bd2225c0bb98a02a63b. 

### Setup
```{r message = FALSE}
# Attach Packages
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(here)
library(tidyverse)
library(janitor)
library(lubridate)
library(kableExtra)
```

## Section 1
### Read in the Data
```{r}
# Read in urchins.csv data and name "urchins_data"
urchins_data <- read_csv(here("data", "urchins.csv"))
```

### Clean Data using clean_names()
```{r}
urchins_clean <- urchins_data %>% 
  clean_names()
```

### Visualize Data: Compare the distribution sizes between purple and red urchins
```{r}
ggplot(data = urchins_clean, aes(x = size)) +
  geom_histogram(color = "red", fill = "white", bins = 10) +
  facet_wrap(~common_name) + 
  labs(x = "Urchin Body Diameter (cm)", 
       y = "Count (#)",
       title = "Purple and Red Urchin Sizes (WILSON)",
       caption = "Data Source: Reed, LTER: Reef:Long-term experiment:
       Kelop removal: Urchin size frequency distribution") + 
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"))
```

**Figure 1: Purple and Red Urchin Sizes** Figure 1 compares the distribution of body diameters (cm) between purple and red urchins. Purple urchin body diameters (cm) have a lower tendency, smaller distribution, and spread than Red Urchins. 

### Descriptive Statistics Table: Create a finalized summary table that contains mean, median, standard deviation, and sample size for purple and red urchins
```{r}
# Create a finalized summary table using kableEXTRA
urchins_summary <- urchins_clean %>% 
  group_by(common_name) %>% 
  summarize(mean = mean(size, na.rm = TRUE),
            median = median(size, na.rm = TRUE),
            sd = sd(size, na.rm = TRUE),
            sample_size = n()) %>% 
  kable(caption = "Table 1: Urchin Body Diameters (cm) Summary Statistics. This table compares red urchins (n=27,408) and purple urchins (n=31,805) body diameters (cm) by mean, median, standard deviation, and sample size",
        col.names = c("Urchin Name",
                      "Mean (cm)",
                      "Median (cm)",
                      "Standard Deviation (cm)",
                      "Sample Size"),
        digits = 1) %>% 
  kable_minimal(html_font = "Georgia", font_size = 12, full_width = FALSE) %>% 
  kable_material(c("striped", "hover", "condensed"))

urchins_summary
```

**Conclusion:** The purple urchins have a mean value of 4.5 (cm), which is 2.3 (cm) less than the red urchin's mean value. This a 40.7% difference between mean values. The purple urchins also have a smaller central tendency and spread; their standard deviation is 1.2(cm), which is 0.7 (cm) less than the red urchin's standard deviation. This is a 45.2% difference between standard deviation. 

## Section 2

### Wrangle Data: Create a data subset that only includes purple urchins. Change dates to mdy format and change site names to full descriptions 
```{r}
purple_urchins <- urchins_clean %>% 
  filter(common_name == "Purple Urchin") %>% 
  mutate(date_new = mdy(date)) %>% 
  mutate(year_new = year(date_new)) %>% 
  mutate(site_full = case_when(
    site == "NAPL" ~ "Naples Reef",
    site == "MOHK" ~ "Mohawk Reef",
    site == "IVEE" ~ "Isla Vista Reef",
    site == "AQUE" ~ "Arroyo Quemado Reef",
    site == "CARP" ~ "Carpinteria Reef"))
```

### Descriptive Statistics Table: Store a summary table containing total number of purple urchins observed by site and year
```{r}
purple_urchin_summary <- purple_urchins %>% 
  group_by(site_full, year_new) %>% 
  summarize(sample_size = n())

purple_urchin_summary
```

### Visualize Data: Compare purple urchin counts at each site from 2008 - 2020
```{r}
ggplot(data = purple_urchin_summary, 
       aes(x = year_new, y = sample_size, color = site_full)) +
  geom_line(size = 0.5) +
  scale_x_continuous(breaks = c(2008, 2010, 2012, 2014, 2016, 2018, 2020)) +
  labs(x = "Year",
       y = "Count (#)",
       title = "Total Purple Urchin Count by Site (2008 -2020) (WILSON)",
       fill = "Site",
       caption = "Data Source: Reed, LTER: Reef:Long-term experiment:
       Kelop removal: Urchin size frequency distribution",
       color = "Site") +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"))
```

**Figure 2: Total Purple Urchin Count by Site** Figure 2 compares the number of purple urchins in five sites between 2008 - 2010. The sites included are; Arroyo Quemado Reef, Carpinteria Reef, Isla Vista Reef, Mohawk Reef, and Naples Reef. All of the sites show an increase in purple urchin count in 2008 followed by an overall decline in purple urchin count from 2009 to 2020. 

**Analysis:** There is a possibility that this graph accurately reflects changes in the purple urchin population. However, you would need to take the sample size into account to determine the accuracy. In general, the larger the sample size, the higher the confidence level. We could apply a t test to determine if there is a significant difference between the sample mean and the population mean. 

The changes in the number of purple urchins observed could be explained by an inaccurate representation of data because of the small sample size and/or recorded inaccurately due to human error. 


## End Task 3
